---
title: "Pileup_Graph"
output: html_notebook
---
Set working directory and resolve imports
```{r}
setwd("~/Documents/ND5_pileup_analysis")
library(ggplot2)
library(dplyr)
library(patchwork)
```
Import File & Transform Data
```{r}
pileup_data <- read.table(file = "ND5_pileup_mutant.csv",sep = ",",header = TRUE)
pileup_data_subset1 <- pileup_data[pileup_data$POS >= 13708,]
pileup_data_subset_mut <- pileup_data_subset1[pileup_data_subset1$POS <= 13712,]
pileup_data_subset_mut$POS_CAT <- as.character(pileup_data_subset_mut$POS)
pileup_data_subset_mut$condition <- "mutant"
write.csv(pileup_data_subset_mut, file = "ND5_mut_subset_m1A_5pos.csv")
```

```{r}
pileup_data1 <- read.table(file = "ND5_pileup_wildtype.csv",sep = ",",header = TRUE)
pileup_data_subset_a <- pileup_data1[pileup_data1$POS >= 13708,]
pileup_data_subset_wt <- pileup_data_subset_a[pileup_data_subset_a$POS <= 13712,]
pileup_data_subset_wt$POS_CAT <- as.character(pileup_data_subset_wt$POS)
pileup_data_subset_wt$condition <- "wildtype"
write.csv(pileup_data_subset_wt, file = "ND5_wt_subset_m1A_5pos.csv")
```
```{r}
alldata <- bind_rows(pileup_data_subset_mut,pileup_data_subset_wt)
```
Create Graphs
```{r}
alldata_plot <- ggplot() +
  geom_boxplot(data=alldata, aes(x=POS_CAT,y=FREQ2,fill=condition), width=.75) + 
  geom_point(data=alldata,aes(x=POS_CAT,y=FREQ2,fill=condition),position=position_dodge2(.75)) +
  ggtitle("ND5 Misincorporation") +
  xlab("Position") +
  ylab("Misincorporation Frequency") +
  ylim(0,.025) +
  theme_classic()
alldata_plot
ggsave("ND5_five_position_comparision.pdf",dpi=800)


