---
title: "R Notebook"
output: html_notebook
author: 'Christina Fitzsimmons'
date: 'Last updated 2022-05-04'
---
# Introduction
This notebook compares and plots GO-terms from metascape analysis and reviGO collapse.

# Upstream Processing
RNA-seq data from Gencode v35 library 1 was processed in DEseq2 and these gene names were supplied to metascape. In metascape, we performed custom analysis, and selected all categories to maximize information we learned. The data are grouped in the following way:

1. Set 1 is padj < 0.05 and l2fc > 1 or l2fc < 1 -  these are named:
    + 2022.05.03_l2fc1_repeat_metascape_upregulated 
    + 2022.05.03_l2fc1_repeat_metascape_downregulated

2. Set 2 is padj < 0.05 and l2fc > 0 or l2fc < 0 -  these are named:
    + 2022.05.03_padj005_repeat_metascape_upregulated 
    + 2022.05.03_padj005_repeat_metascape_downregulated

3. Set 3 is all genes with padj <0.05 – named:
    + 2022.05.03_padj005_all_metascape

```{r}
library(tidyverse)
library(ggpubr)
library(httr)
library(stringi)

theme_set(theme_bw())

Set1_down  <- read_csv("2021.02_analysis_gencodeV35/2022.05.03_l2fc1_repeat_metascape_downregulated/Enrichment_GO/_FINAL_GO.csv")
Set1_up <- read_csv("2021.02_analysis_gencodeV35/2022.05.03_l2fc1_repeat_metascape_upregulated/Enrichment_GO/_FINAL_GO.csv")


Set2_down <-read_csv("2021.02_analysis_gencodeV35/2022.05.03_padj005_repeat_metascape_downregulated/Enrichment_GO/_FINAL_GO.csv")
Set2_up <- read_csv("2021.02_analysis_gencodeV35/2022.05.03_padj005_metascape_upregulated/Enrichment_GO/_FINAL_GO.csv")

Set3_all <- read_csv("2021.02_analysis_gencodeV35/2022.05.03_padj005_all_metascape/Enrichment_GO/_FINAL_GO.csv")



```
# Filter the lists for GO-Biological Process terms
reviGO doesn't work with terms from other categories. 

```{r}

# Filter GO BP terms only
Set1_down_GOBP <- Set1_down %>%
  dplyr::filter(Category == "GO Biological Processes")

Set1_up_GOBP <- Set1_up %>%
  dplyr::filter(Category == "GO Biological Processes")


Set2_down_GOBP <- Set2_down %>%
  dplyr::filter(Category == "GO Biological Processes")

Set2_up_GOBP <- Set2_up %>%
  dplyr::filter(Category == "GO Biological Processes")

Set3_all_GOBP <- Set3_all %>%
  dplyr::filter(Category == "GO Biological Processes") 


# write out these files for reviGO web interface
write_csv(Set1_down_GOBP, file = "2021.02_analysis_gencodeV35/2022.05.03_Set1_down_GOBP.csv", col_names = TRUE)
write_csv(Set1_up_GOBP, file ="2021.02_analysis_gencodeV35/2022.05.03_Set1_up_GOBP.csv", col_names = TRUE )

write_csv(Set2_down_GOBP, file = "2021.02_analysis_gencodeV35/2022.05.03_Set2_down_GOBP.csv", col_names = TRUE)
write_csv(Set2_up_GOBP, file ="2021.02_analysis_gencodeV35/2022.05.03_Set2_up_GOBP.csv", col_names = TRUE )

write_csv(Set3_all_GOBP, file = "2021.02_analysis_gencodeV35/2022.05.03_Set3_all_GOBP.csv", col_names = TRUE)

```
# Make plots of the reviGO data
```{r}
Set1_down_revigo  <- read_tsv("2021.02_analysis_gencodeV35/2022.05.03_l2fc1_repeat_metascape_downregulated/reviGO_collapse_set1_down/2022.05.03_revigo_collapse_filter07_set1_down.tsv")

set1_down2 <- read_tsv("2021.02_analysis_gencodeV35/2022.05.03_l2fc1_repeat_metascape_downregulated/reviGO_collapse_set1_down/2022.05.04_revigo_collapse_set1_down_filter07.tsv")

Set1_up_revigo <- read_tsv("2021.02_analysis_gencodeV35/2022.05.03_l2fc1_repeat_metascape_upregulated/reviGO/2022.05.03_revigo_collapse_filter07_set1_up.tsv")


Set2_down_revigo <-read_tsv("2021.02_analysis_gencodeV35/2022.05.03_padj005_repeat_metascape_downregulated/reviGO/2022.05.03_revigo_collapse_filter07_set2_down.tsv")
Set2_up_revigo <- read_tsv("2021.02_analysis_gencodeV35/2022.05.03_padj005_metascape_upregulated/reviGO/2022.05.03_revigo_collapse_filter07_set2_up.tsv")

Set3_07_revigo <- read_tsv("2021.02_analysis_gencodeV35/2022.05.03_padj005_all_metascape/reviGO/2022.05.03_revigo_collapse_filter07_Set3_all.tsv")
Set3_05_revigo <- read_tsv("2021.02_analysis_gencodeV35/2022.05.03_padj005_all_metascape/reviGO/2022.05.03_revigo_collapse_filter05_set3_all.tsv")
Set3_04_revigo <- read_tsv("2021.02_analysis_gencodeV35/2022.05.03_padj005_all_metascape/reviGO/2022.05.03_revigo_collapse_filter04_set3_all.tsv")
```
# plotting top 10 terms for set 3
```{r}

set3_07_top10 <- Set3_07_revigo %>%
  dplyr::filter(Eliminated == FALSE) %>%
  dplyr::arrange(Value) %>% 
  dplyr::slice(1:10) %>% # select the first 10 GO-terms
  dplyr::mutate("log10P" = -1* Value) %>%
  dplyr::mutate(TermID = factor(TermID, TermID)) # This ranks the GO-terms in my order, rather than alphabetically. 

set3_05_top10 <- Set3_05_revigo %>%
  dplyr::filter(Eliminated == FALSE) %>%
  dplyr::arrange(Value) %>% 
  dplyr::slice(1:10) %>% # select the first 10 GO-terms
  dplyr::mutate("log10P" = -1* Value) %>%
  dplyr::mutate(TermID = factor(TermID, TermID)) # This ranks the GO-terms in my order, rather than alphabetically. 

set3_04_top10 <- Set3_04_revigo %>%
  dplyr::filter(Eliminated == FALSE) %>%
  dplyr::arrange(Value) %>% 
  dplyr::slice(1:10) %>% # select the first 10 GO-terms
  dplyr::mutate("log10P" = -1* Value) %>%
  dplyr::mutate(TermID = factor(TermID, TermID)) # This ranks the GO-terms in my order, rather than alphabetically. 

```
# Plotting the Set 3 bar plots
```{r}
# ReviGO plot
set3_07_plot <- ggplot(set3_07_top10 , aes(x = log10P, y =TermID)) +
  geom_col(fill = "gray80") +
  geom_text(aes(label=Name), 
            hjust = 0, x=0, size = 3) +
  labs(title = "GO Biological Pathways: Genes with padj < 0.05",
       subtitle = "ReviGO collapse 0.7",
       x = "-Log10P", y = "") +
  scale_y_discrete(limits = rev(set3_07_top10$TermID))
set3_07_plot 

set3_05_plot <- ggplot(set3_05_top10 , aes(x = log10P, y =TermID)) +
  geom_col(fill = "gray80") +
  geom_text(aes(label=Name), 
            hjust = 0, x=0, size = 3) +
  labs(title = "GO Biological Pathways: Genes with padj < 0.05",
       subtitle = "ReviGO collapse 0.5",
       x = "-Log10P", y = "") +
  scale_y_discrete(limits = rev(set3_05_top10$TermID))
set3_05_plot


set3_04_plot <- ggplot(set3_04_top10 , aes(x = log10P, y =TermID)) +
  geom_col(fill = "gray80") +
  geom_text(aes(label=Name), 
            hjust = 0, x=0, size = 3) +
  labs(title = "GO Biological Pathways: Genes with padj < 0.05",
       subtitle = "ReviGO collapse 0.4",
       x = "-Log10P", y = "") +
  scale_y_discrete(limits = rev(set3_04_top10$TermID))
set3_04_plot

# Exporting the figures
pdf("2022.05.04_Set3_padj05_revigo07_barplot.pdf", 
   useDingbats=FALSE,
   width=5, height=4)
set3_07_plot
dev.off()

pdf("2022.05.04_Set3_padj05_revigo05_barplot.pdf", 
   useDingbats=FALSE,
   width=5, height=4)
set3_05_plot
dev.off()


pdf("2022.05.04_Set3_padj05_revigo04_barplot.pdf", 
   useDingbats=FALSE,
   width=5, height=4)
set3_04_plot
dev.off()

```
# Revigo Plots of Set 1 (l2fC > ± 1 and padj <0.05)
```{r}

Set1_down_10 <- Set1_down_revigo %>%
  dplyr::filter(Eliminated == FALSE) %>%
  dplyr::arrange(Value) %>% 
  dplyr::slice(1:10) %>% # select the first 10 GO-terms
  dplyr::mutate("log10P" = -1* Value) %>%
  dplyr::mutate(direction = 'downregulated') %>%
  dplyr::mutate(TermID = factor(TermID, TermID)) # This ranks the GO-terms in my order, rather than alphabetically. 

Set1_up_10 <- Set1_up_revigo %>%
  dplyr::filter(Eliminated == FALSE) %>%
  dplyr::arrange(Value) %>% 
  dplyr::slice(1:10) %>% # select the first 10 GO-terms
  dplyr::mutate("log10P" = -1* Value) %>%
  dplyr::mutate(direction = 'upregulated') %>%
  dplyr::mutate(TermID = factor(TermID, TermID)) # This ranks the GO-terms in my order, rather than alphabetically. 

set1_down2_10 <- set1_down2 %>%
  dplyr::filter(Eliminated == FALSE) %>%
  dplyr::arrange(Value) %>% 
  dplyr::slice(1:10) %>% # select the first 10 GO-terms
  dplyr::mutate("log10P" = -1* Value) %>%
  dplyr::mutate(direction = 'upregulated') %>%
  dplyr::mutate(TermID = factor(TermID, TermID)) # This ranks the GO-terms in my order, rather than alphabetically. 


Set1_combined <- bind_rows(Set1_down_10, Set1_up_10)
set1_filtered <- dplyr::filter(Set1_combined, log10P > 5)

```
# plotting set 1 data
```{r}

# ReviGO plot
set1_down_plot <- ggplot(Set1_down_10 , aes(x = log10P, y =TermID)) +
  geom_col(fill = "gray80") +
  geom_text(aes(label=Name), 
            hjust = 0, x=0, size = 3) +
  labs(title = "GO Biological Pathways",
       subtitle = "Set1: downregulated",
       x = "-Log10P", y = "") +
  scale_y_discrete(limits = rev(Set1_down_10$TermID))
set1_down_plot

set1_up_plot <- ggplot(Set1_up_10 , aes(x = log10P, y =TermID)) +
  geom_col(fill = "gray80") +
  geom_text(aes(label=Name), 
            hjust = 0, x=0, size = 3) +
  labs(title = "GO Biological Pathways",
       subtitle = "Set1: upregulated",
       x = "-Log10P", y = "") +
  scale_y_discrete(limits = rev(Set1_up_10$TermID))
set1_up_plot


set1_all_plot <- ggplot(Set1_combined , aes(x = log10P, y =TermID, fill = direction)) +
  geom_col(show.legend = FALSE, stat = "TermID") +
  facet_grid(direction ~ ., scales = "free") +
  geom_text(aes(label=Name), 
            hjust = 0, x=0, size = 3) +
  labs(title = "GO-BP Up and Downregulated in Mutant cells",
       subtitle = "Set 1: reviGO collapse 0.7",
       x = "-Log10P", y = "") +
  scale_y_discrete(limits = rev)
set1_all_plot


set1_filter_plot <- ggplot(set1_filtered , aes(x = log10P, y =TermID, fill = direction)) +
  geom_col()+
  geom_text(aes(label=Name), 
            hjust = 0, x=0, size = 3) +
  labs(title = "GO-BP Up and Downregulated in Mutant cells",
       subtitle = "Set 1: reviGO collapse 0.7",
       x = "-Log10P", y = "") +
  scale_y_discrete(limits = rev)
set1_filter_plot


```
# Export the Set1 figures
```{r}
# Exporting the figures
pdf("2022.05.04_Set1_up_l2fc1_padj05_revigo07_barplot.pdf", 
   useDingbats=FALSE,
   width=5, height=4)
set1_up_plot
dev.off()

pdf("2022.05.04_Set1_down_l2fc1_padj05_revigo07_barplot.pdf", 
   useDingbats=FALSE,
   width=5, height=4)
set1_down_plot
dev.off()


pdf("2022.05.04_Set1_combined_l2fc1_padj05_revigo07_facet_barplot.pdf", 
   useDingbats=FALSE,
   width=5, height=4)
set1_all_plot
dev.off()

pdf("2022.05.04_Set1_combined_l2fc1_padj05_revigo07_filter_barplot.pdf", 
   useDingbats=FALSE,
   width=5, height=4)
set1_filter_plot
dev.off()

```
# Set 2

```{r}
Set2_down_10 <- Set2_down_revigo %>%
  dplyr::filter(Eliminated == FALSE) %>%
  dplyr::arrange(Value) %>% 
  dplyr::slice(1:10) %>% # select the first 10 GO-terms
  dplyr::mutate("log10P" = -1* Value) %>%
  dplyr::mutate(direction = 'downregulated') %>%
  dplyr::mutate(TermID = factor(TermID, TermID)) # This ranks the GO-terms in my order, rather than alphabetically. 

Set2_up_10 <- Set2_up_revigo %>%
  dplyr::filter(Eliminated == FALSE) %>%
  dplyr::arrange(Value) %>% 
  dplyr::slice(1:10) %>% # select the first 10 GO-terms
  dplyr::mutate("log10P" = -1* Value) %>%
  dplyr::mutate(direction = 'upregulated') %>%
  dplyr::mutate(TermID = factor(TermID, TermID)) # This ranks the GO-terms in my order, rather than alphabetically. 


Set2_combined <- bind_rows(Set2_down_10, Set2_up_10)
set2_filtered <- dplyr::filter(Set2_combined, log10P > 5)
```
# Plotting set 2
```{r}
# ReviGO plot
set2_down_plot <- ggplot(Set2_down_10 , aes(x = log10P, y =TermID)) +
  geom_col(fill = "gray80") +
  geom_text(aes(label=Name), 
            hjust = 0, x=0, size = 3) +
  labs(title = "GO Biological Pathways",
       subtitle = "Set2: downregulated",
       x = "-Log10P", y = "") +
  scale_y_discrete(limits = rev(Set2_down_10$TermID))
set1_down_plot

set2_up_plot <- ggplot(Set2_up_10 , aes(x = log10P, y =TermID)) +
  geom_col(fill = "gray80") +
  geom_text(aes(label=Name), 
            hjust = 0, x=0, size = 3) +
  labs(title = "GO Biological Pathways",
       subtitle = "Set2: upregulated",
       x = "-Log10P", y = "") +
  scale_y_discrete(limits = rev(Set2_up_10$TermID))
set2_up_plot


set2_all_plot <- ggplot(Set2_combined , aes(x = log10P, y =TermID, fill = direction)) +
  geom_col(show.legend = FALSE, stat = "TermID") +
  facet_grid(direction ~ ., scales = "free") +
  geom_text(aes(label=Name), 
            hjust = 0, x=0, size = 3) +
  labs(title = "GO-BP Up and Downregulated in Mutant cells",
       subtitle = "Set 2: reviGO collapse 0.7",
       x = "-Log10P", y = "") +
  scale_y_discrete(limits = rev)
set2_all_plot


set2_filter_plot <- ggplot(set2_filtered , aes(x = log10P, y =TermID, fill = direction)) +
  geom_col()+
  geom_text(aes(label=Name), 
            hjust = 0, x=0, size = 3) +
  labs(title = "GO-BP Up and Downregulated in Mutant cells",
       subtitle = "Set 2: reviGO collapse 0.7",
       x = "-Log10P", y = "") +
  scale_y_discrete(limits = rev)
set2_filter_plot
```
# Export the figures
```{r}
# Exporting the figures
pdf("2022.05.04_Set2_up_l2fc1_padj05_revigo07_barplot.pdf", 
   useDingbats=FALSE,
   width=5, height=4)
set2_up_plot
dev.off()

pdf("2022.05.04_Set2_down_l2fc1_padj05_revigo07_barplot.pdf", 
   useDingbats=FALSE,
   width=5, height=4)
set2_down_plot
dev.off()


pdf("2022.05.04_Set2_combined_l2fc1_padj05_revigo07_facet_barplot.pdf", 
   useDingbats=FALSE,
   width=5, height=4)
set2_all_plot
dev.off()

pdf("2022.05.04_Set2_combined_l2fc1_padj05_revigo07_filter_barplot.pdf", 
   useDingbats=FALSE,
   width=5, height=4)
set2_filter_plot
dev.off()
```

