---
title: "Analysis of Riboseq data"
output: html_notebook
author: 'Christina Fitzsimmons'
date: 'Last updated 202-06-15'
---

## 2022-06-15
## Generate STAR index for MANE transcriptome
Please use the following publication to cite MANE: Morales, J., Pujar, S., Loveland, J.E. et al. A joint NCBI and EMBL-EBI transcript set for clinical genomics and research. Nature (2022). DOI: 10.1038/s41586-022-04558-8

https://ftp.ncbi.nlm.nih.gov/refseq/MANE/MANE_human/release_0.9/README.txt
```{bash, eval=FALSE}
# make sure all gtf and fasta files are unzipped or this mode will fail!
set -o pipefail
set -e

function fail {
    echo "$@" >&2
    exit 1
    }

module load samtools/1.11         || fail "could not load samtools module"
module load STAR/2.7.6a          || fail "could not load STAR module"

cd /data/BatistaLab_NGS/annotations/MANE
mkdir -p repeat_Index_STAR2.7.6a

STAR \
  --runThreadN $SLURM_CPUS_PER_TASK \
  --runMode genomeGenerate \
  --genomeDir repeat_Index_STAR2.7.6a \
  --genomeFastaFiles /data/BatistaLab_NGS/annotations/MANE/MANE.GRCh38.v0.9.select_refseq_rna.fna \
  --sjdbGTFfile /data/BatistaLab_NGS/annotations/MANE/MANE.GRCh38.v0.9.select_refseq_genomic.gtf \
  --sjdbOverhang 50

```

Fatal INPUT FILE error, no valid exon lines in the GTF file: /data/BatistaLab_NGS/annotations/MANE/MANE.GRCh38.v0.9.select_refseq_genomic.gtf
Solution: check the formatting of the GTF file. One likely cause is the difference in chromosome naming between GTF and FASTA file.

```{bash}

```
## 2022-06-16
Problem was genomic gtf with transcriptomic fasta file. Removing the GTF and overhang lines fixes the problem. 

# Mapping to MANE genome
`sbatch --array=0 --partition=ccr --mem=40g --cpus-per-task=8 --mail-type=END,FAIL --gres=lscratch:50 STAR2.7.6_mapping_MANE.sh`
```{bash, eval =FALSE}
#! /bin/bash

export TMPDIR=/lscratch/$SLURM_JOB_ID
set -o pipefail
set -e

function fail {
    echo "$@" >&2
    exit 1
    }

module load samtools/1.11         || fail "could not load samtools module"
module load STAR/2.7.6a         || fail "could not load STAR module"

cd /data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped || fail "no such directory"
mkdir -p riboseq3_MANEv0.9
GENOME=/data/BatistaLab_NGS/annotations/MANE/MANE_Index_STAR2.7.6a/

declare -A INPUTS
INPUTS[0]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/splitFastq/map_repeats/BC2rUnmapped.out.mate1"
INPUTS[1]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/splitFastq/map_repeats/BC4rUnmapped.out.mate1"
INPUTS[2]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/splitFastq/map_repeats/BC5rUnmapped.out.mate1"
INPUTS[3]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/splitFastq/map_repeats/BC6rUnmapped.out.mate1"
INPUTS[4]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/splitFastq/map_repeats/BC7rUnmapped.out.mate1"
INPUTS[5]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/splitFastq/map_repeats/BC12rUnmapped.out.mate1"
INPUTS[6]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/splitFastq/map_repeats/BC13rUnmapped.out.mate1"
INPUTS[7]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/splitFastq/map_repeats/BC14rUnmapped.out.mate1"
INPUTS[8]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/splitFastq/map_repeats/BC15rUnmapped.out.mate1"
INPUTS[9]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/splitFastq/map_repeats/BC16rUnmapped.out.mate1"
INPUTS[10]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/splitFastq/map_repeats/BC18rUnmapped.out.mate1"
INPUTS[11]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/splitFastq/map_repeats/BC19rUnmapped.out.mate1"
INPUT=${INPUTS[$SLURM_ARRAY_TASK_ID]}

declare -A OUTNAMES
OUTNAMES[0]=WTinput1
OUTNAMES[1]=WTinput2
OUTNAMES[2]=WTinput3
OUTNAMES[3]=MUTinput1
OUTNAMES[4]=MUTinput2
OUTNAMES[5]=MUTinput3
OUTNAMES[6]=WTrpf1
OUTNAMES[7]=WTrpf2
OUTNAMES[8]=WTrpf3
OUTNAMES[9]=MUTrpf1
OUTNAMES[10]=MUTrpf2
OUTNAMES[11]=MUTrpf3
OUTNAME=${OUTNAMES[$SLURM_ARRAY_TASK_ID]}


STAR \
    --runThreadN $SLURM_CPUS_PER_TASK \
    --genomeDir $GENOME \
    --readFilesIn $INPUT \
    --outSAMtype BAM SortedByCoordinate \
    --outWigType wiggle \
    --outTmpDir=/lscratch/$SLURM_JOB_ID/STARtmp \
    --outReadsUnmapped Fastx \
    --outFileNamePrefix riboseq3_MANEv0.9/$OUTNAME \
    –-outFilterType BySJout \
    --outFilterMultimapNmax 10 \
    --outFilterMismatchNmax 999 \
    --outFilterMismatchNoverLmax 0.04

```
## Use samtools to index 
According to plastid, this is the next step. This creates a .bai file for each sorted bam file
`samtools index WTrpf2Aligned.sortedByCoord.out.bam `

## 1. Generate regions of interest
```{bash, eval=FALSE}

metagene generate cdsstop --landmark cds_stop --annotation_files /data/BatistaLab_NGS/annotations/MANE/MANE.GRCh38.v0.9.select_ensembl_genomic.gtf --downstream 50 --upstream 100

metagene generate cdsstart --landmark cds_start --annotation_files /data/BatistaLab_NGS/annotations/MANE/MANE.GRCh38.v0.9.select_ensembl_genomic.gtf --downstream 100 --upstream 50

```
# Creating a new conda environment
```{bash}

# To use the newly installed conda source an init file. Do this each time you are going to work with your environments.
source /data/$USER/conda/etc/profile.d/conda.sh

# Activate the base environment (upon which all other environments are built)
conda activate base

which python # check the conda python is being used instead of the biowulf system python
conda update conda # update the conda package manager
conda clean --all --yes # removes all unused packages

```
### 3.1.2 Creating a plastid-specific environment
The next step is creating a python environment specifically for Plastid
```{bash}
# Generate a new environment using python 3.8
# Make sure whatever version of python you select for your new environment matches the base conda python--or they won't be able to communicate

conda create -name plastid python=3.8

# Once created, activate the environment to install package dependencies
conda activate plastid

# install with EITHER pip3
pip3 install plastid

```
## 2022-06-22
Trying to use mamba and pip as per HPC suggestions for installing plastid environment

```{bash}

# activate the plastid conda environment and install mamba
conda activate plastid
conda install -c conda-forge mamba

# Install plastid and dependencies
mamba install -c bioconda plastid

```
 Custom conda environment made with mamba worked! now onto genome generation and counting
```{bash}
# gene position generation:
cs generate --annotation_files genes.gtf --annotation_format GTF2 cs_generate_genes

# counting-- all BAM files must first be indexed with samtools
cs count MANE_position_gene.positions --count_files /data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_MANEv0.9/OnePass/BC02Aligned.sortedByCoord.out.bam --fiveprime --countfile_format BAM --min_length 25 --max_length 34 BC2_Mane_counts

```
 All the count files are empty--ENST vs ENSG counting is the issue. 
 1) try this with gencode mapped files
 2) guydosh pipeline is specific for MANE
 3) make or find ENST gtf file for MANE
 
## 2022-06-23
```{bash}
# Counting gencode against gencode gene.positions



# Counting gencode against MANE gene.positions
cs count /data/BatistaLab_NGS/Riboseq3/Nextseq_2000/plastid_metagene_counts/MANE/MANE_position_gene.positions --count_files /data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC02Aligned.sortedByCoord.out.bam --fiveprime --countfile_format BAM --min_length 25 --max_length 34 BC02MANEcount

```
# Are they similar?
```{r}
library(tidyverse)

BC02_gencode <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC02count.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, utr5_reads) %>%
  dplyr::rename(utr5_gencode = utr5_reads)

BC02_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC02MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, utr5_reads) %>%
  dplyr::rename(utr5_MANE = utr5_reads)

counts <- dplyr::inner_join(BC02_gencode, BC02_MANE, by = c("region"))

scatter1 <- ggplot(counts, aes(x = utr5_gencode, y = utr5_MANE)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  xlim(0,500) +
  ylim(0,500)

scatter1
```
# counting all the files at one time
Do I need an offset? Only for the metagene? Will ask in lab...
```{bash}
#!/bin/bash
# Christina Fitzsimmons
set -e
set -o pipefail

source /data/fitzsimmonscm/conda/etc/profile.d/conda.sh

declare -A INPUTS
INPUTS[0]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC02Aligned.sortedByCoord.out.bam"
INPUTS[1]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC04Aligned.sortedByCoord.out.bam"
INPUTS[2]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC05Aligned.sortedByCoord.out.bam"
INPUTS[3]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC06Aligned.sortedByCoord.out.bam"
INPUTS[4]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC07Aligned.sortedByCoord.out.bam"
INPUTS[5]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC12Aligned.sortedByCoord.out.bam"
INPUTS[6]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC13Aligned.sortedByCoord.out.bam"
INPUTS[7]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC14Aligned.sortedByCoord.out.bam"
INPUTS[8]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC15Aligned.sortedByCoord.out.bam"
INPUTS[9]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC16Aligned.sortedByCoord.out.bam"
INPUTS[10]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC18Aligned.sortedByCoord.out.bam"
INPUTS[11]="/data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC19Aligned.sortedByCoord.out.bam"
INPUT=${INPUTS[$SLURM_ARRAY_TASK_ID]}

declare -A OUTNAMES
OUTNAMES[0]=BC02MANEcount
OUTNAMES[1]=BC04MANEcount
OUTNAMES[2]=BC05MANEcount
OUTNAMES[3]=BC06MANEcount
OUTNAMES[4]=BC07MANEcount
OUTNAMES[5]=BC12MANEcount
OUTNAMES[6]=BC13MANEcount
OUTNAMES[7]=BC14MANEcount
OUTNAMES[8]=BC15MANEcount
OUTNAMES[9]=BC16MANEcount
OUTNAMES[10]=BC18MANEcount
OUTNAMES[11]=BC19MANEcount
OUTNAME=${OUTNAMES[$SLURM_ARRAY_TASK_ID]}

conda activate plastid

cs count /data/BatistaLab_NGS/Riboseq3/Nextseq_2000/plastid_metagene_counts/MANE/MANE_position_gene.positions \
--count_files $INPUT \
--fiveprime \
--countfile_format BAM \
--min_length 25 \
--max_length 34 \
$OUTNAME
```

# trying to plot the metagene
```{bash}
source /data/fitzsimmonscm/conda/etc/profile.d/conda.sh
conda activate plastid

metagene count cdsstart_rois.txt --count_files /data/BatistaLab_NGS/Riboseq3/Nextseq_2000/mapped/riboseq3_gencode/OnePass/BC16Aligned.sortedByCoord.out.bam --threeprime --normalize_over "-50" 100 --use_mean --keep -v --countfile_format BAM BC16_test

```
This wasn't really working. Will ask Guydosh lab for help running. Likely user error. 


# DATA ANALYSIS
```{r, Load Libraries}
library(DESeq2)
library(tidyverse)
library(circlize)
library("RColorBrewer")
library("gplots")
library(ggrepel)
library(ggpubr)
library(gprofiler2)
library(plotly)
library(corrplot)
library(ggcorrplot)
library(pheatmap)
```
# Section 2: Gencode Mapped, MANE gene.positions, plastid counted
Here, we're going to analyze the plastid counts

```{r}
# Wildtype Data
BC02_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC02MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC02 = cds_reads)

BC04_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC04MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC04 = cds_reads)

BC05_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC05MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC05 = cds_reads)

BC06_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC06MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC06 = cds_reads)

BC07_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC07MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC07 = cds_reads)

BC12_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC12MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC12 = cds_reads)


# Mutant Data
BC13_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC13MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC13 = cds_reads)

BC14_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC14MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC14 = cds_reads)

BC15_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC15MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC15 = cds_reads)

BC16_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC16MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC16 = cds_reads)

BC18_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC18MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC18 = cds_reads)

BC19_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC19MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC19 = cds_reads)
```
# merging the dataframes together and bringing in the metadata
```{r}
#put all data frames into list
df_list <- list(BC02_MANE, BC04_MANE, BC05_MANE, BC06_MANE, BC07_MANE, BC12_MANE, BC13_MANE, BC14_MANE, BC15_MANE, BC16_MANE, BC18_MANE, BC19_MANE)

#merge all data frames in list
df_list %>% reduce(full_join, by='region')

#convert back to df format
df_table <- as.data.frame(df_list) %>%
  dplyr::select(region, BC02, BC04, BC05, BC06, BC07, BC12, BC13, BC14, BC15, BC16, BC18, BC19) %>%
  column_to_rownames(var = "region")

# import the metadata
col_data <- read.delim(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/Analysis/riboseq_metadata.txt", header = TRUE, row.names = 1)

```
# Section 2: Gencode Mapped, MANE gene.positions, plastid counted
Here, we're going to analyze the plastid counts

```{r}
# Wildtype Data
BC02_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC02MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC02 = cds_reads)

BC04_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC04MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC04 = cds_reads)

BC05_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC05MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC05 = cds_reads)

BC06_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC06MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC06 = cds_reads)

BC07_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC07MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC07 = cds_reads)

BC12_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC12MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC12 = cds_reads)


# Mutant Data
BC13_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC13MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC13 = cds_reads)

BC14_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC14MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC14 = cds_reads)

BC15_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC15MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC15 = cds_reads)

BC16_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC16MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC16 = cds_reads)

BC18_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC18MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC18 = cds_reads)

BC19_MANE <- read.table(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/RAW/plastid_gencode_OnePass/BC19MANEcount.txt" , skip = 18, header = TRUE, sep = "\t") %>%
  dplyr::select(region, cds_reads) %>%
  dplyr::rename(BC19 = cds_reads)
```
# merging the dataframes together and bringing in the metadata
```{r}
# put all data frames into list
df_list <- list(BC02_MANE, BC04_MANE, BC05_MANE, BC06_MANE, BC07_MANE, BC12_MANE, BC13_MANE, BC14_MANE, BC15_MANE, BC16_MANE, BC18_MANE, BC19_MANE)

# merge all data frames in list
df_list %>% reduce(full_join, by='region')

# convert back to df format
df_table <- as.data.frame(df_list) %>%
  dplyr::select(region, BC02, BC04, BC05, BC06, BC07, BC12, BC13, BC14, BC15, BC16, BC18, BC19) %>%
  column_to_rownames(var = "region")

# import the metadata
col_data <- read.delim(file = "/Users/fitzsimmonscm/Documents/Projects_and_Data/Batista_Lab/FH_SDHB_Project/Rprojects/UOK_manuscript/Ribosome_profiling/Analysis/riboseq_metadata.txt", header = TRUE, row.names = 1)


# creating DESeq dds object from treatments and count data
dds <- DESeqDataSetFromMatrix(countData = countdata,
                              colData = col_data,
                              design = ~ celltype + condition + celltype:condition)

# touching up dds object by filtering out any rows with count sums less than 10
nrow(dds)
dds <- dds[rowSums(counts(dds)) >10,]
nrow(dds)
counts(dds)

```
# Running the analysis
```{r}

# Running deseq2 analysis using the multi-factor model
dds <- DESeq(dds)
res <- results(dds)
summary(res)

# what comparisons is Deseq making?
resultsNames(dds)

```
# volcano plot of data
```{r}
res_stats <- as.data.frame(res) %>%
  na.omit


# make volcano plot
testvolcano <- res_stats %>%
  mutate("logDESeqP" = -1*log10(padj)) %>%
  mutate(
    "direction" = case_when(
      log2FoldChange <= -1 & padj < 0.05 ~ "downreg",
      log2FoldChange >= 1 & padj < 0.05 ~ "upreg",
      TRUE ~ "no_change")) %>%
  dplyr::mutate(direction = factor(direction, levels = c("upreg", "downreg", "no_change")))

ribo_scatter <- ggplot(testvolcano, aes(x = log2FoldChange, y = logDESeqP, color=direction)) +
  geom_point() +
  theme_bw() +
  geom_vline(xintercept = -1, linetype="dashed") + geom_vline(xintercept = 1, linetype = "dashed") +
  geom_hline(yintercept = 1.30, linetype = "dashed") +
  scale_color_manual(values=c("black", "black", "black")) +
  #scale_color_manual(values=c("red", "blue", "#999999")) +
  labs(x = "Log2FoldChange (MUT/WT)", y = "-Log10(P)", title = "Multifactor Analysis of CDS counts", subtitle = "celltype + condition + celltype:condition") 
ribo_scatter

#gprofiler query of gene names
annotated_volcano <- testvolcano %>%
  rownames_to_column("ENSG") %>%
  separate(ENSG, into = c("ENSG_ID", NA), sep = "[.]")

results_genes = gconvert(annotated_volcano$ENSG_ID, organism = "hsapiens",
			     target = "ENTREZGENE_ACC", filter_na = FALSE)
head(results_genes)


# add the gene names to the ENSG ID info
annotated_volcano = merge(annotated_volcano,
		   results_genes[,c("input", "name")],
		      by.x = "ENSG_ID", by.y = "input")

# write out information into a CSV file
write.csv(res_stats, file = "2022.07_MUT_over_WT_RPF_multifactor_differential_expression.csv")
```
